---
title: "Geog533 Lab 10"
author: "Your  name"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r}
if(!require(UScensus2010)){install.packages("UScensus2010")}
library(UScensus2010)
if(!require(UScensus2010county)) {install.county("osx")}
library(UScensus2010county)
data(new_york.county10)
shp_NY <- new_york.county10
plot(shp_NY)
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r}
if(!require(tmap)) {install.packages("tmap")}
library(tmap)
qtm(shp_NY,fill = "NAME10",text = "NAME10",text.size = 0.5
    ,title = "NY county")
```


### Question 1(c)	
How many counties in New York State?
```{r}
df_NY <- shp_NY@data
nrow(df_NY)
### there are 62 counties in New York State.
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
df_NY[df_NY$NAME10 == "Broome",]$fips
#The 3-digit fips code of Broome County is 007.
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r}
pop_NY <- df_NY$P0010001
su <- summary(pop_NY)
total <- sum(pop_NY)
if(!require(e1071)){install.packages("e1071")}
library(e1071)
sk <- skewness(pop_NY)
min <- su[[1]]
max <- su[[6]]
median <- su[[3]]
mean <- su[[4]]
result <- cbind(total,min,max,mean,median,sk)
rownames(result) <- "output"
df1 <-data.frame(result)
if(!require(knitr)){install.packages("knitr")}
library(knitr)
kable(df1)
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(pop_NY)
```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)


### Question 2(a)	
Plot a map of New York census tracts using the plot function.
```{r}
if(!require(UScensus2010tract)) {install.tract("osx")}
library(UScensus2010tract)
data("new_york.tract10")
shp_NYtr <- new_york.tract10
plot(shp_NYtr)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
df_NYtr <- shp_NYtr@data
pop_NYtr <- df_NYtr$P0010001
sum(pop_NYtr)
# The total population based on census tracts is 19378102
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
broome <- df_NYtr$county == "007"
plot(shp_NYtr[broome,])
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
df_broome <- df_NYtr[df_NYtr$county == "007",]
pop_broome <- df_broome$P0010001
sum(pop_broome)
# The total population of Broome County is 200600
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(pop_broome)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r, message=FALSE, warning=FALSE}
df_broome_sel5 <- head(df_NYtr,5)
#df_broome_sel51 <- df_NYtr[1:5,]
df_broome_sel5$cpop <- sum(pop_broome)
df_broome_sel5$ratio <- df_broome_sel5$P0010001/df_broome_sel5$cpop
df_broome_sel5
shp_broome <- shp_NYtr[broome,]
shp_broome_sel5 <- shp_broome[1:5,]
shp_broome_sel5@data <- df_broome_sel5
if(!require(rgdal)) install.packages("rgdal")
library(rgdal)
#writeOGR(shp_broome,dsn = "output",layer = "shp_broome",driver = "ESRI Shapefile", overwrite_layer = TRUE)
writeOGR(shp_broome_sel5,dsn = "output",layer = "shp_broome_sel5",driver = "ESRI Shapefile", overwrite_layer = TRUE)

```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)
```{r, message=FALSE, warning=FALSE}
if(!require(ncdf4)) {install.packages("ncdf4")}
library(ncdf4)
library(raster)
```

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
ndvi_mb <- brick("NDVI.nc")
ndvi_mb
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(ndvi_mb)
ncol(ndvi_mb)
ncell(ndvi_mb)
nbands(ndvi_mb)
extent(ndvi_mb)
res(ndvi_mb)
bbox(ndvi_mb)
projection(ndvi_mb)

```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
mean_ndvi <- mean(ndvi_mb)
plot(mean_ndvi)
max_ndvi <- max(ndvi_mb)
plot(max_ndvi)
writeRaster(x = mean_ndvi,filename = "mean_ndvi.tif", overwrite = TRUE)
writeRaster(x = max_ndvi,filename = "max_ndvi.tif",overwrite = TRUE)
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
ndvi_mb_2001 <- ndvi_mb[[13:24]]
plot(ndvi_mb_2001)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r, message=FALSE, warning=FALSE}
hist(ndvi_mb_2001)
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}
ndvi_mb_2000_7 <- ndvi_mb[[7]]
plot(ndvi_mb_2000_7)
#values <- click(ndvi_mb, n=1, xy=TRUE) 
values <- click(ndvi_mb, n=1, xy=FALSE)
values2 <- ndvi_mb[50,90]
plot(as.vector(values2), type="b")
```

