---
title: "Web Mapping"
author: "Dr. Qiusheng Wu"
output: 
  html_notebook: 
    toc: yes
    toc_float: yes
---

## Useful websites

 * [htmlwidget for R](http://www.htmlwidgets.org/showcase_leaflet.html)
 
 * [Leaflet for R](https://rstudio.github.io/leaflet/)
 
 * [Leaflet basemap](http://leaflet-extras.github.io/leaflet-providers/preview/)
 
 * [Leaflet Application 1](http://wetlands.io/maps/hometown.html)
 
 * [Leaflet Application 2](http://wetlands.io/maps/connectivity.html)
 
 
## Leaflet

### Installation

To install this R package, run this command at your R prompt:
```{r, message=FALSE, warning=FALSE}
if(!require(leaflet)) {install.packages("leaflet")}
```

Once installed, you can use this package at the R console, within R Markdown documents.


### Basic Usage

You create a Leaflet map with these basic steps:

1. Create a map widget by calling leaflet().

2. Add layers (i.e., features) to the map by using layer functions (e.g. addTiles, addMarkers, addPolygons) to modify the map widget.

3. Repeat step 2 as desired.

4. Print the map widget to display it.

Here’s a basic example:

```{r}
m <- leaflet(width = "100%")
m <- addTiles(m)  # Add default OpenStreetMap map tiles
m <- addMarkers(m, lng=-75.968342, lat=42.089026, popup="Binghamton University")
m  # Print the map
```


### Pipe operator

```{r}
m <- leaflet(width = "100%") %>%
  addTiles() %>%
  addMarkers(m, lng=-75.968342, lat=42.089026, popup="Binghamton University")
m
```


### Basemaps
The easiest way to add tiles is by calling `addTiles()` with no arguments; by default, OpenStreetMap tiles are used.

Alternatively, many popular free third-party basemaps can be added using the `addProviderTiles()` function, which is implemented using the leaflet-providers plugin. See [here](http://leaflet-extras.github.io/leaflet-providers/preview/index.html) for the complete set. 
```{r}
m <- leaflet(width = "100%") %>%
  addTiles() %>%
  addProviderTiles(provider = "Esri.WorldStreetMap",group = "World StreetMap") %>%
  addProviderTiles(provider = "Esri.WorldImagery",group = "World Imagery") %>%
  addProviderTiles(provider = "NASAGIBS.ViirsEarthAtNight2012",group = "Nighttime Imagery") %>%
  addMarkers(m, lng=-75.968342, lat=42.089026, popup="Binghamton University") %>%
  addLayersControl(
    baseGroups = c("OSM (default)","World StreetMap", "World Imagery", "Nighttime Imagery"),
    options = layersControlOptions(collapsed = FALSE))
m
```

### Markers

Use markers to call out points on the map. Marker locations are expressed in latitude/longitude coordinates, and can either appear as icons or as circles. You can use `SpatialPoints`, `SpatialPointsDataFrame`, or a two-column numeric matrix (first column is longitude, second column is latitude) as data objects.

You can also use a data frame with latitude and longitude columns. You can explicitly tell the marker function which columns contain the coorinate data (e.g. `addMarker(lng = ~Longitude, lat = ~Latitude)`), or let the function look for columns named `lat/latitude` and `lon/lng/long/longitude` (case insensitive).

#### Icon Markers
Icon markers are added using the `addMarker` function. Their default appearance is a dropped pin. As with most layer functions, the `popup` argument can be used to add a message to be displayed on click.

```{r}
library(knitr)
df <- quakes 
kable(head(df,n = 10)) # display the first 10 records

leaflet(data = df, width = "100%") %>% 
  addTiles() %>%
  addMarkers(~long, ~lat,popup = ~as.character(mag))
```

#### Marker Clusters

```{r}
df <- quakes 
leaflet(data = df, width = "100%") %>% 
  addTiles() %>%
  addMarkers(~long, ~lat,popup = ~as.character(mag),clusterOptions = markerClusterOptions())
  
```


## Dygraphs

<http://rstudio.github.io/dygraphs>

Dygraphs provides rich facilities for charting time-series data in R and includes support for many interactive features including series/point highlighting, zooming, and panning.

### Installation
```{r}
if(!require(dygraphs)) {install.packages("dygraphs")}
```

### Basic Usage

```{r}
library(dygraphs)
library(knitr)
df <- nhtemp
kable(head(df,n = 10)) ## display the first 10 records
dygraph(nhtemp, main = "New Haven Temperatures", width = "100%") %>% 
  dyRangeSelector(dateWindow = c("1920-01-01", "1960-01-01"))

```

### Example

<http://wetlands.io/maps/PPR_NCDC.html>
