---
title: "Geog533 Lab 8"
author: "Your name"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
ssr1 <- 1.7
sst1 <- 2.3
n <- 24
ssm1 <- sst1 - ssr1
df11 <- 1
df21 <- n - 2
dft1 <- n - 1
msm1 <- ssm1/1
msr1 <- ssr1/(n-2)
f1 <- msm1/msr1
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|       `r ssm1`          |   `r df11`                 |        `r msm1`        |      `r f1`           |
| Residual |       1.7       |      `r df21`              |      `r msr1`          |                 |
| Total    |       2.3       |       `r dft1`             |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
a <- 0.46
b <- 0.19
x1 <- 50
y1 <- b*x1 + a
y1
#predicted number of weekly shopping trips is 9.96
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?

Answer: when one unit increases in annual income, the weekly shopping trip frequency will increase 0.19.


### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
f_critical <- qf(0.95,df1 = 1,df2 = 22)
if(f1 > f_critical){
  print("we reject the non hypothesis")
}else {
  print("we can not reject non hypothesis")
}
#regression cofficient significantly different form zero
```



### Question 1(e)
What is the value of the correlation coefficient?
```{r}
R <- sqrt(ssm1/sst1)
R
# the correlation coefficient value is 0.5107539
```





## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
sn <- c(36,78,11,45)
el <- c(400,800,200,675)
df <- data.frame(el,sn)
m <- lm(df$sn~df$el)
summary(m)
m$coefficients
# intercept and slope are not significant
```


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
result <- summary(m)
residual_sd <- result$sigma
residual_sd

```


### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r, message=FALSE, warning=FALSE}
result <- summary(m)
result
if(result$coefficients[1,4] > 0.05){
  print("we can not reject non hypothesis")
}
c_f <- qf(0.95,df1 = 1,df2 = 2)
if(result$fstatistic < c_f){
  print("regression coefficient is equal to zero")
}
```


### Question 2(d)
Find the value of $r^2$.
```{r}
result$r.squared
#the value of r^2 is 0.889949
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
library(knitr)
ob_v <- df$sn
pre_v <- m$fitted.values
res_v <- m$residuals
df_tab <- data.frame(ob_v,pre_v,res_v)
kable(df_tab)
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
result_a <- anova(m)
result_a
df1 <- result_a$Df[1]
df2 <- result_a[2,1]
ssm <- result_a$`Sum Sq`[1]
ssr <- result_a$`Sum Sq`[2]
sst <- ssm + ssr
dft <- df1 + df2
msm <- result_a$`Mean Sq`[1]
msr <- result_a$`Mean Sq`[2]
f <- msm/msr
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|     `r ssm`            |        `r df1`            |        `r msm`        |     `r f`            |
| Residual |     `r ssr`            |         `r df2`           |        `r msr`        |                 |
| Total    |     `r sst`            |         `r dft`           |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
plot(df)
abline(m,col = "red",lwd = 2)
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
df3 <- read.csv(url, header = T)
m <- lm(df3$SalePrice~df3$LotSize)
m
result3 <- summary(m)
result3
f_c <- qf(0.95,df1 = 1,df2 = nrow(df3)-2)
if(result3$fstatistic[1] > f_c){
  print("we reject non hypothesis, regression coefficient is significant.")
}else{
  print("we can not reject non hypothesis")
}
plot(df3$LotSize,df3$SalePrice)
abline(m,col = "red",lwd = 2)
```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
df4 <- read.csv(url,header = T)
m <- lm(df4$price~df4$bedrooms)
m
result4a <- summary(m)
result4a
f_c <- qf(0.95,df1 = 1,df2 = nrow(df4)-2)
if(result4a$fstatistic[1] > f_c){
  print("we reject non hypothesis, regression coefficient is significant.")
}else{
  print("we can not reject non hypothesis")
}
plot(df4$bedrooms,df4$price)
abline(m,col = "red",lwd = 2)
```


### Question 4(b) 
(b) repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
m <- lm(df4$price~df4$bathrooms)
m
result4b <- summary(m)
result4b
if(result4b$fstatistic[1] > f_c){
  print("we reject non hypothesis, regression coefficient is significant.")
}else{
  print("we can not reject non hypothesis")
}
plot(df4$bathrooms,df4$price)
abline(m,col = "red",lwd = 2)
```


## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r}
if(!require(leaflet)) {install.packages("leaflet")}
library(leaflet)
df <- read.csv(file = "https://xujianzi.github.io/data/hometown1.csv",header = T)
content <- paste(sep = "<br/>",
                 "<b><a href='https://xujianzi.github.io/data/WechatIMG75.jpeg'>Jian Xu</a></b>",
                 "Yufeng International ",
                 "Tianxian road"
)
m <- leaflet(data = df,width = "100%") %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(~longitude,~latitude,popup = ~fullname,label = ~image) %>%
  addPopups(113.8660792,30.9191275,content,options = popupOptions(closeButton = F))
m  # Print the map
```

