---
title: "Using R to do Raster analysis"
author: "Xu Jian"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---



## Description
The data are landsat8 June to Augest one day image from 2013 to 2017.
Choosing Binghamton as the Stuady Area
Implement 1.Mask
          2.Calculate NDVI
          3.Extract Vegetation
          4.Calculate Dense Vegetation Area
If you want to analyze other Raster data, you just need to reset the list of the
filepath and change the "n"

## Librarys
Using "raster","rgdal","UScensus2010","UScensus2010county","knitr"
```{r, message=FALSE, warning=FALSE}
if(!require(raster)){install.packages("raster")}
library(raster)
if(!require(rgdal)){install.packages("rgdal")}
library(rgdal)
if(!require(UScensus2010)) {install.packages("UScensus2010")}
library(UScensus2010)
if(!require(UScensus2010county)) {install.county("osx")}
library(UScensus2010county)
if(!require(knitr)){install.packages("knitr")}
library(knitr)

```


## Define the function
### NDVI function
Define the function to calculate NDVI from land8 image
```{r}
ndvCalc <- function(x) {
  ndvi <- (x[[4]] - x[[3]]) / (x[[4]] + x[[3]])
  return(ndvi)
}
ndvOver <- function(x, y) {
  ndvi <- (y - x) / (x + y)
  return(ndvi)
}
```

### Vegetation function
Define the function to get the vegetation area and plot it
```{r}
getVegetation <- function(x){
  x[x < 0.2] <- NA
  return(x)
}
```

### Dense Vegetation function
Define the function to calculate the dense vegetation area
```{r}
getArea <- function(x){
  return(sum(x[x>=0.6])*res(x)[1]^2)
}
```

## Set work directory and list file
```{r, message=FALSE, warning=FALSE}
listf <- list()
listf[1] <- "G:/R/FinalProject/data/land8_2017"
listf[2] <- "G:/R/FinalProject/data/lan8_2016"
listf[3] <- "G:/R/FinalProject/data/land8_2015"
listf[4] <- "G:/R/FinalProject/data/land8_2014"
listf[5] <- "G:/R/FinalProject/data/land8_2013"
listp <- list()
bandcom <- list()
n <- 5
for(i in 1:n){
  setwd(listf[[i]])
  getwd()
  listp[[i]] <- list.files(pattern= "*.TIF",full.names=TRUE)
  bandcom[[i]] <- stack(listp[[i]][4:9])
}
#### show landsat8 image in 2017 as an example
bandcom[[1]]
names(bandcom[[1]])
plot(bandcom[[1]])
```


## Reprojection Rasters
reproject all the Raster layers into one Projection(including extents)
```{r}
newbandcom <- list()
for(i in 1:5){
  newbandcom[[i]] <-  projectRaster(bandcom[[i]],bandcom[[1]])
}

```

## Shapefile
Get the broome county shapefile 
```{r}
data("new_york.county10")
shp_NY <- new_york.county10
df_NY <- shp_NY@data
broome <- df_NY$NAME10 == "Broome"
shp_Broome <- shp_NY[broome,]
plot(shp_Broome)
shp_Broome
```

## Reproject shapefile
Reproject shapefile to landsat8 images
The coordinate of shapefile is WGS84 and landsat8 imges is from UTM series 
```{r}
shp_BroomeUTM <- spTransform(shp_Broome, CRS(proj4string(bandcom[[1]])))
shp_BroomeUTM 
```
## Raster Analysis
### Mask
```{r, message=FALSE, warning=FALSE}
band_broome <- list()
for(i in 1:n){
  band_broome[[i]] <- mask(bandcom[[i]],shp_BroomeUTM)
}
#### use 2016 images as an example
plot(band_broome[[2]],ext = shp_BroomeUTM)
names(band_broome[[2]])
```

### NDVI
Calculate NDVI in Broome using overlay()
```{r, message=FALSE, warning=FALSE}
ndvi <- list()
for(i in 1:n){
  ndvi[[i]] <- overlay(x=band_broome[[i]][[3]], y=band_broome[[i]][[4]], fun=ndvOver)
}
```

### Vegetation
Extract the NDVI > 0.2 which is sparse vegetation and dense vegetation
```{r, message=FALSE, warning=FALSE}
vegetation <- list()
for(i in 1:n){
  vegetation[[i]] <- getVegetation(ndvi[[i]])
}
```

### Dense Vegetation Area
Calculate the area of dense vegetation 
```{r, message=FALSE, warning=FALSE}
Dveg_Area <- list()
for(i in 1:n){
  Dveg_Area[i] <- getArea(ndvi[[i]])
}
area <- as.numeric(Dveg_Area)
year <- c("2017","2016","2015","2014","2013")
df <- data.frame(year,area)
kable(df)
plot(df,type = "b")
```

### Display ndvi plot and hist
The layer1 to layer3 is 2017 to 2013
```{r, message=FALSE, warning=FALSE}
for(i in 1:n){
  plot(ndvi[[i]],ext = shp_BroomeUTM)
hist(ndvi[[i]])
}
```

### Display vegetation
The layer1 to layer3 is 2017 to 2015
The layer1 to layer2 is 2014 to 2013
```{r, message=FALSE, warning=FALSE}
# vege_show1 <- brick(vegetation[[1]],vegetation[[2]],vegetation[[3]])
# plot(vege_show1,ext = shp_BroomeUTM)
# hist(vege_show1)
# vege_show2 <- brick(vegetation[[4]],vegetation[[5]])
# plot(vege_show2,ext = shp_BroomeUTM)
# hist(vege_show2)
for(i in 1:n){
  plot(vegetation[[i]],ext = shp_BroomeUTM)
  hist(vegetation[[i]])
}
```
