---
title: "Xu_Jian_lab3"
author: "Jian Xu"
date: "9/28/2017"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Question 1
### Part 1 
#### Randomly select 40 cars as one dataset and remaining as the other dataset
```{r}
library(MASS)
df_lab3 <- Cars93
index <- nrow(df_lab3)
set.seed(100)
training <- sample(1:index, size = 40, replace = FALSE)
whole <- as.numeric(row.names(df_lab3)) 
remain <- whole[-training]
df_t <- df_lab3[training,]
df_r <- df_lab3[remain,]
write.csv(df_t, file = "df_t.csv")
write.csv(df_r, file = "df_r.csv")
```
### Part 2
#### Find the number of USA and non-USA cars in the training dataset
```{r}
USA_cars <- df_t[df_t$Origin == "USA",]
Num_USA <- nrow(USA_cars)
non_USA_cas <- df_t[df_t$Origin == "non-USA",]
Num_non_USA <- nrow(non_USA_cas)
result_1_2 <-cbind(Num_USA, Num_non_USA)
rownames(result_1_2) <- "Numbers"
result_1_2
```
## Question 2
```{r}
set.seed(210)
S_trial <- sample(c("boy","girl"), size = 10, replace = T, prob = c(0.515, 0.485))
set.seed(210000)
L_trial <- sample(c("boy","girl"), size = 10000, replace = T, prob = c(0.515, 0.485))
result_2 <- cbind(table(S_trial), table(L_trial))
colnames(result_2) <- c("small_trial", "large_trial")
result_2
```

## Question 3 use Binomial Distribution
### Part 1 
#### Density function ##boy=X  P(X = 3)
```{r}
dbinom(3, size = 10, prob = 0.515)
```
### Part 2
#### Distribution fuction ## P(X >= 8) 
It can translated to caculate the girls'
```{r}
g_p2 <- pbinom(2, size = 10, prob = 0.485)
g_p2
```
### Part 3
#### Density function plot
```{r}
n <- 10
x <- 0:n
y <- dbinom(x, size = 10, prob = 0.515)
plot(x, y, type ="b", pch = 16, col="blue")
title(main = list("The Possibility of Numbers of boys in 10 births", col = "red", font = 3))
```
#### Cumulative probability function plot
```{r}
n <- 10
x <- 0:n
y <- pbinom(x, 10, prob = 0.515)
plot(x, y, type = "b", pch = 16, col = "green")
title(main = list("The Cumulative Possibility", col ="red", font = 3))
```

## Question 4 use Poisson Distribution
### Part 1
##### Exactly four cars arriving in ten seconds
```{r}
dpois(4, lambda = 5)
```
### Part 2 
##### More than five cars arriving in ten seconds
```{r}
ppois(5, lambda = 5, lower.tail = F)
```
### Part 3
```{r}
Ncar_max <- 20
Ncar <- 0:Ncar_max
P_car <- dpois(Ncar, lambda = 5)
plot(Ncar, P_car, type = "b", pch = 16, col = "red")
title(main = list("The Possibility of Numbers of arriving cars", 
                  col = "red", font = 3))
```



