---
title: "Xu_jian_lab4"
author: "Jian Xu"
date: "October 5, 2017"
output: 
  html_notebook:
    toc: yes
    toc_float: yes
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Question 1
### Part a
```{r}
answer_1_a <- pnorm(20, 32, 7) + dnorm(20, 32, 7)
print(answer_1_a)
```
The answer is **0.05635001** 

Plot the density function
```{r}
cord.x <- c(11, seq(11, 20, 0.01), 20)
cord.y <- c(0,dnorm(seq(11, 20, 0.01), 32,7),0)
curve(dnorm(x, 32, 7), xlim = c(11, 53), 
      main = "A team scoring 20 points or fewer")
polygon(cord.x, cord.y, col = "red")
```

### Part b
```{r}
answer_1_b <- pnorm(35, 32, 7, lower.tail = F)
print(answer_1_b)
```
The answer is **0.3341176** 

Plot the density function
```{r}
cord.x <- c(35, seq(35,53,0.01), 53)
cord.y <- c(0, dnorm(seq(35, 53, 0.01), 32, 7), 0)
curve(dnorm(x, 32, 7), xlim = c(11, 53), 
      main = "A team scoring more than 35 points")
polygon(cord.x, cord.y, col = "blue")
```

### Part c
```{r}
answer_1_c <- pnorm(40,32,7) - pnorm(20, 32, 7)
print(answer_1_c)
```
The answer is **0.8302129** 

Plot the density function
```{r}
cord.x <- c(20, seq(20,40,0.01), 40)
cord.y <- c(0, dnorm(seq(20, 40, 0.01), 32, 7), 0)
curve(dnorm(x, 32, 7), xlim = c(11, 53), 
      main = "A team scoring between 20 and 40 points")
polygon(cord.x, cord.y, col = "yellow")
```


## Question 2
### Part a
```{r}
answer_2_a <- pexp(3,0.1)
print(answer_2_a)
```
The answer is **0.2591818**

Plot the density function
```{r}
curve(dexp(x,0.1), xlim = c(0, 50), lwd = 2,
      main = "Fewer than 3 comments")
cord.x <- c(0, seq(0, 3, 0.01), 3)
cord.y <- c(0,dexp(seq(0,3,0.01), 0.1),0)
polygon(cord.x, cord.y, col = "purple")
```

### Part b
```{r}
answer_2_b <- pexp(20,0.1,F)
print(answer_2_b)
```
The answer is **0.1353353**

Plot the density function
```{r}
curve(dexp(x,0.1), xlim = c(0, 50), lwd = 2,
      main = "More than 20 comments")
cord.x <- c(20, seq(20, 50, 0.01), 50)
cord.y <- c(0,dexp(seq(20,50,0.01), 0.1),0)
polygon(cord.x, cord.y, col = "pink")
```

### Part c
```{r}
answer_2_c <- pexp(10, 0.1) - pexp(5, 0.1)
print(answer_2_c)
```
The answer is **0.2386512**

Plot the density function
```{r}
curve(dexp(x,0.1), xlim = c(0, 50), lwd = 2,
      main = "Between 5 and 10 comments")
cord.x <- c(5, seq(5, 10, 0.01), 10)
cord.y <- c(0,dexp(seq(5,10,0.01), 0.1),0)
polygon(cord.x, cord.y, col = "black")
```


## Question 3
```{r, message=FALSE, warning=FALSE}
library(raster)
```
### Part a
30*30 raster based on random values from uniform distribution
```{r}
ras <- raster(nrows = 30, ncols = 30, xmn = 0,
              xmx = 30, ymn = 0, ymx = 30)
ras 
set.seed(100)
ras[] <- runif(ncell(ras))
plot(ras)
hist(ras)
```

### Part b
30*30 raster based on random values from normal distribution
```{r}
ras2 <- raster(nrows = 30, ncols = 30, xmn =0,
               xmx = 30, ymn = 0, ymx = 30)
ras2
set.seed(99)
ras2[] <- rnorm(ncell(ras2))
plot(ras2)
hist(ras2)
```

### Part c
cell-by-cell addition of first raster and second raster
```{r}
ras3 <- ras + ras2
plot(ras3)
hist(ras3)
```

### Part d
Calculate the mean value of the third raster

reclassify the third raster(cell value 0,1)
```{r, message=FALSE, warning=FALSE}
if(!require(rgdal)) install.packages("rgdal")
library(rgdal)
```

```{r}
mean_R3 <- cellStats(ras3, "mean")
print(mean_R3)
ras3[ras3 > mean_R3] <- 1
ras3[ras3 <= mean_R3] <- 0
plot(ras3)
writeRaster(ras3, filename = "ras3_reclassify.tiff", overwrite = TRUE)
```

